Query Optimization for Dynamic Imputation

### 2017-05-11 paper/experiments.tex
paper: pass over experiments

\subsection{Queries}
We collect a set of queries (\Cref{fig:queries}) that we think are interesting to plan.
We believe that they could reasonably be written by a user in the course of data analysis.

@@

Our first set of queries is on the CDC data (\Cref{fig:queries-cdc}).
\Cref{q1} calculates the average cuff size for individuals based on their income data, with a constraint on height.
\Cref{q2} compares creatine levels for individuals with low, medium, and high incomes and above a certain weight.
\Cref{q3} extracts the average blood lead levels for children under 6 years of age.
\Cref{q4} calculates the average systolic blood pressure, by gender, for subjects with a body mass index indicating obesity. 
\Cref{q5} calculates the average waist circumference for subjects above a certain height and weight.

Our second set of queries is on the freeCodeCamp data (\Cref{fig:queries-fcc}).
\Cref{q6} calculates the average income for higher-income survey participants, grouped by their bootcamp attendance.
\Cref{q7} estimates the average commute time of women from the United States who participated.
\Cref{q8} calculates the average amount of student debt based on school degree for survey participants who have student debt.
\Cref{q9} joins the freeCodeCamp data with a table provided by the World Bank which summarizes GDP per-capita across various countries~\cite{worldbank-data}.
The query calculates the average GDP per-capita by grouping 18+ year old participants who attended bootcamp versus those who did not.

\begin{table*}
  \centering
  \input{queries}
\end{table*}

@@

\Cref{fig:runtimes} shows a summary of the performance results. The quality-optimized
queries are an order-of-magnitude faster than imputing on the base tables. We can get
another order-of-magnitude speedup when optimizing for performance. This performance
differential means it is feasible
to explore multiple imputations operations (including more expensive operators) when using
\ProjectName{}, in contrast to the traditional approach of base table imputation.

@@

In the median across a variety of queries and choices of $\alpha$, planning took
\planningmediantime{} ms, 
constituting \planningruntimepercent{} percent of total runtime. In all cases, the optimizer
returned a query plan within \planningmaxtime{} ms, with times similar across values of $\alpha$.

@@

We explored the practical implications of our planning algorithm by constructing a series of queries with increasing number of joins.
We used the CDC tables and evaluated a combination of self-joins and joins across reference tables. Planning times for queries
involving 1 to 5 joins, which although limited is a practical value for real world exploratory queries, take less than a second.
For queries with 7 joins, planning time increases substantially to 29 seconds, on average. We introduced an optimization that reduces
the plan space by maintaining only approximations of the Pareto set for queries involving a large number of joins.
With this optimization, the same set of queries have an average planning time of 16 seconds (2 seconds SE).

>>>>

\subsection{Queries}
We collect a set of queries (\Cref{tbl:queries}) that we think are interesting to plan.
We believe that they could reasonably be written by a user in the course of data analysis.

@@

The first set of queries is on the CDC data.
\Cref{q1} calculates the average cuff size for individuals based on their income data, with a constraint on height.
\Cref{q2} compares creatine levels for individuals with low, medium, and high incomes and above a certain weight.
\Cref{q3} extracts the average blood lead levels for children under 6 years of age.
\Cref{q4} calculates the average systolic blood pressure, by gender, for subjects with a body mass index indicating obesity. 
\Cref{q5} calculates the average waist circumference for subjects above a certain height and weight.

The second set of queries is on the freeCodeCamp data.
\Cref{q6} calculates the average income for higher-income survey participants, grouped by their bootcamp attendance.
\Cref{q7} estimates the average commute time of women from the United States who participated.
\Cref{q8} calculates the average amount of student debt based on school degree for survey participants who have student debt.
\Cref{q9} joins the freeCodeCamp data with a table provided by the World Bank which summarizes GDP per-capita across various countries~\cite{worldbank-data}.
The query calculates the average GDP per-capita by grouping 18+ year old participants who attended bootcamp versus those who did not.

\begin{table*}
  \centering
  \input{queries}
  \vspace{1ex}
  \caption{Queries used in our experiments.}\label{tbl:queries}
\end{table*}

@@

\Cref{fig:runtimes} shows the performance results. The quality-optimized
queries are an order-of-magnitude faster than imputing on the base tables. We get
another order-of-magnitude speedup when optimizing for performance. This performance
differential means it is feasible
to explore multiple imputation operations (including more expensive operators) when using
\ProjectName{}, in contrast to the traditional approach of base table imputation.

@@

@@

Planning times are reasonable, with a median planning time of \planningmediantime{} ms. over the benchmark queries, constituting \planningruntimepercent{}\% of total runtime.
In all cases, the optimizer returned a query plan within \planningmaxtime{} ms.
The choice of $\alpha$ has no impact on planning times, because query selection is done after collecting the Pareto frontier.

However, \ProjectName{}'s planning algorithm is still exponential in the number of joins in the plan, and the use of Pareto sets exacerbates this exponential growth.
We explored the limits our planning algorithm by constructing a series of queries with increasing number of joins.
We used the CDC tables and evaluated a combination of self-joins and joins across reference tables.
Planning times for queries involving 1--5 joins, which is a practical value for real world exploratory queries, take less than a second.
For queries with 7 joins, planning time increases substantially to 29 seconds, on average.
We introduced an optimization that reduces the plan space by maintaining an approximation of the Pareto set for queries involving a large number of joins.
With this optimization, the same set of queries have an average planning time of 16 seconds (2 seconds SE).
